---
title: "PML - Course Project"
author: "Simon C"
date: "6 April 2017"
output: html_document
---
#Predicting Personal Activity
In this assignment we will predict the classification of 5 different activities recorded by 6 participants in a study.  

##Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).
##Data
The training data for this project are available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)
  
The test data are available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)
  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir= '~/Desktop/coursera/Assignments/PML')
```

```{r, message=FALSE, warning=FALSE}
#set up libraries
library(caret)
library(ggplot2)
library(rpart)
library(randomForest)
```

We will first load the data.  
```{r}
#load data
trainingdata <- read.csv('./data/pml-training.csv')
problem <- read.csv('./data/pml-testing.csv')
```
The training data consists of 19,622 observations of 160 observations.  The test data consists of 20 observations of 53 variables.  

The variables missing from the test data have missing values in the training set and so we will remove them. The class response in the test set will need to be predicted.   We will split the training data into training and validation sets.   

```{r}
#remove X and username
trainingdata <- subset(trainingdata, select=!(names(trainingdata) %in% c('X','user_name','raw_timestamp_part_1','raw_timestamp_part_2','cvtd_timestamp','new_window','num_window')))

problem <- subset(problem, select=!(names(problem) %in% c('X','user_name','raw_timestamp_part_1','raw_timestamp_part_2','cvtd_timestamp','new_window','num_window')))

a <- sapply(problem,function(x){ sum(is.na(x))} )
problem <- subset(problem, select=!(names(problem) %in% names(a[a==20])))

trainingdata <- subset(trainingdata, select=(names(trainingdata) %in% c(names(problem),'classe')))


#split trainingdata into train and test
set.seed(123321)
intrain <- createDataPartition(trainingdata$classe, p=0.7, list=FALSE)

train <- trainingdata[intrain,]
test <- trainingdata[-intrain,]

```

##Prediction
We will first fit a single tree to the data to understand the based accuracy.  
```{r cachedChunk1, cache=TRUE}
modelRPART <- rpart(classe ~ ., data = train, method="class")
confusionMatrix(train$classe,predict(modelRPART, newdata=train, type="class")) 
```

The single tree was only 73% accurate.  

We will now fit a Random Forest to the training data.  Limiting growth to 20 trees creates a model that is 99.99% accurate on the training set.  

```{r cachedChunk2, cache=TRUE}
#Fit RF

modelRF <- randomForest(classe ~ ., data=train, ntree=20)

confusionMatrix(train$classe,predict(modelRF, newdata=train))                        

```
The model also is 99% accurate on the validation set.

```{r}
confusionMatrix(test$classe,predict(modelRF, newdata=test))                   
```
##Importance
The importance plot shows that 8 variables have the most impact on the classification.

```{r}
varImpPlot(modelRF, type=2, col=1, cex=0.8)
```

##Prediction
Finally we will predict the classes of the 20 test observations.
```{r}
predict(modelRF, newdata = problem)

```

